Tidy data structure to support exploration and modeling of temporal-context data

### 2018-11-01 main.Rmd
added duplicated data entry

@@

The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics (reference). It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format. This kind of event describes each flight scheduled for departure at a time point in its local time zone. But every single flight can be uniquely identified by its flight number and origin, from a passenger's point of view. For example, flight DL771 connects multiple origins and destinations rather than a specific route, shown in Figure \@ref(fig:dl771). This dataset is intrinsically heterogeneous, encoding in numbers, strings, and date-times.

@@

@@

The tsibble framework, as expected, incorporates this type of data, without the loss of data richness and heterogeneity. To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", columns `flight` and `origin` as "key" via `id(flight, origin)`. As a result of event data, this data is irregularly spaced, and hence switching to irregular option is necessary. The program internally validates if the key and index produce the distinct rows, and then sort the key and the index from past to recent. When the tsibble creation is done, the print display is data-oriented and contextually informative, such as dimensions, irregular interval (``r tsibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tsibble::tbl_sum(us_flights)[2]``).

@@

>>>>

It encourage users to look into the data as earlier as possible, examining the data quality issues.

@@

The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics (reference). It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format. This kind of event describes each flight scheduled for departure at a time point in its local time zone. Every single flight would be uniquely identified by the flight number and its scheduled departure time, from a passenger's point of view. In fact, it fails to pass the tsibble hurdle due to duplicates in the original data. An error is immediately raised when attempting to convert this data into a tsibble, and a closer inspection has to be carried out to locate the issue. The **tsibble** package provides the tool to easily find the duplicates in the data with `find_duplicates()`. Below shows the problematic entries.

```{r find-duplicate}
```

The issue is perhaps introduced when updating or entering the data into a system. The same flight is scheduled at exactly the same time, together with the same performance statistics but different flight details, which is very unlikely. Flight NK630 is usually scheduled at 17:45 from Chicago to New York, searching into the whole records. A decision is made on removal of the first row from the duplicated entries before proceeding to the tsibble creation.

<!--

@@

-->

@@

This dataset is intrinsically heterogeneous, encoding in numbers, strings, and date-times. The tsibble framework, as expected, incorporates this type of data, without the loss of data richness and heterogeneity. To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", column `flight_num` as "key" via `id(flight_num)`. As a result of event data, this data is irregularly spaced, and hence switching to irregular option is necessary. The program internally validates if the key and index produce the distinct rows, and then sort the key and the index from past to recent. When the tsibble creation is done, the print display is data-oriented and contextually informative, such as dimensions, irregular interval (``r tsibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tsibble::tbl_sum(us_flights)[2]``).

@@

[Change New York Airports to others to be able to see seasonal effects]
